---
title: "Predicting the correctness of exercise"
author: "Jayson Cheng"
date: "24 May, 2015"
output: html_document
---
# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

# Question definition
We want to, given data from fitness devices, predict if the exercise is performed in the correct or incorrect manner.

# Data 
The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

## Reading the data
We will first read the data sets.

```{r}
orgTrainData <- read.csv('pml-training.csv', na.strings=c("NA",""))
orgTestData <- read.csv('pml-testing.csv', na.strings=c("NA",""))
```

## Preprocessing
Next, we will have to pre-process this data to remove NA values and keep only the useful predictors. Near zero values are also removed.

```{r, warning=FALSE, message=FALSE}
library(caret)
trainClean <- orgTrainData[, colSums(is.na(orgTrainData)) == 0]
trainClean <- trainClean[,-(1:7)]
end <- ncol(trainClean)
trainClean[,-end] <- data.frame(sapply(trainClean[,-end], as.numeric))
nzv <- nearZeroVar(trainClean[, -end], saveMetrics=TRUE)
trainClean <- trainClean[,!as.logical(nzv$nzv)]

testClean <- orgTestData[, sapply(orgTestData, is.numeric)]
```

## Partitioning the data for cross-validation

```{r}
set.seed(12321) 
inTrain <- createDataPartition(trainClean$classe, p=0.75, list = FALSE)
cvTrain <- trainClean[inTrain,]
cvTest <- trainClean[-inTrain,]
```

# Training and prediction on cross-validation data
The random forest model is chosen for this analysis. Prediction is made on the cross-validation test data.

```{r, warning=FALSE, message=FALSE}
library(randomForest) 
model <- randomForest(classe ~ ., data=cvTrain, method="class")
pred <- predict(model, cvTest, type="class")
```

## Testing results with confusionMatrix

```{r}
confusionMatrix(pred, cvTest$classe)
```

An accuracy of ***99.6%*** is achieved with the random forest model, with out-of-sample error rate being ***0.4%***. 

# Prediction on original test data
We proceed to predict the results for the test data with our model.

```{r}
predictOrgTest <- predict(model, testClean, type="class")
predictOrgTest
```
